\chapter*{Abstract}
\label{cha:abstract}

Contours can be understood as the shape or format of a object. In
Music contour can represent a parameter in function of another, like
pitch in function of density or density in function of
amplitude. Contours are important because, as well as pitch sets and
motives, they can help giving coherence to a musical piece.

Theories of contours have been used in areas such as Ear Training and
Analysis, but the systematic use of contours for generation of
compositional material is an issue still lacking literature.

In this thesis I present the piece \obra{} [\opus{Around the
  pomegranate}] and its analysis. This piece, for woodwind quintet,
was composed using combinations of contour operations associated with
parameters such as pitch, tempo, density and texture. In order to
accomplish this task, I did a literature review of contour theories, I
did a mapping of contours to musical elements, I composed studies of
possibilities for experimentation with contours, I develop the Goiaba,
a software to assist in processing contours for composition, and
finally composed the piece \obra{}.

This study helps to advance the state of art of contour theories
through composition contour operations experiments using contour
operations and contributes with new tools to the composition field.
%% Conclude that the
My conclusion is that contours can be used in a systematic way in
musical composition, but we still need further study. Thus this depth
and continuity in development of Goiaba are possible future activities
resulting from this work.

Keywords: Musical composition, Musical contours, Woodwind quintet,
Computer music.
